Challenges and Solutions of Developing and Implementing a Revolutionary Novel Desktop-as-a-Service

### 2024-04-26 paper.tex
Some improvements

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%
\begin{document}
%
\title{Challenges and Solutions of Developing and
	Implementing a Revolutionary Novel Desktop-as-a-Service}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Christian Baun \and 
Johannes Bouché}
% \author{Christian Baun\orcidID{0009-0004-9955-3752} \and
% 	Johannes Bouché\orcidID{TODO}}
%
\authorrunning{C. Baun, J. Bouché}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Faculty of Computer Science and Engineering,\\
Frankfurt University of Applied Sciences,\\
Nibelungenplatz 1, 60318 Frankfurt am Main, Germany
\email{[christianbaun|johannes.bouche]@fb2.fra-uas.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Desktop-as-a-Service (DaaS) is a cloud service category with great potential for improving the everyday work of running and using applications. However, existing DaaS solutions have several drawbacks, which have limited the popularity of this service category in the past. Our novel DaaS solution not only overcomes the limitations of existing solutions but also offers significant benefits. It has the potential to revolutionize the way we use applications, supporting non-modified Linux and Windows applications with just a browser on the client side. This paper describes some of the knowledge we gained during the development because of the challenges we faced and the solutions we found.

@@

This paper discusses challenges during the development and implementation of a novel Desktop-as-a-Service (DaaS) solution that enables the deployment and usage of unmodified Linux and Windows applications in the same way as web applications and can run inside public resources but can also be deployed in a private context. Since all interaction with the DaaS and the applications that reside inside is done via a user's web browser, the users can use any client, no matter what hardware or host operating system it is compromised of. A browser is the only software component required to use the novel DaaS service.

@@

This document is structured as follows. Section~\ref{sec:relatedworkAbschnitt} discusses related work on DaaS solutions from an academic perspective, whereas section~\ref{sec:RecommendDaaSarchitecture} describes the state-of-the-art DaaS architecture we designed and implemented. Section~\ref{sec:AnalysisPossibleComponents} presents the most relevant challenges we faced, analyses our options for handling these, and describes the solution we found. Finally, section~\ref{sec:Conclusions} discusses conclusions, and section~\ref{sec:NextSteps} includes directions for future work.

@@

Celesti et al.~\cite{celesti2016improving} implemented in 2016 a DaaS using OpenStack~\cite{sefraoui2012openstack} and analyzed the characteristics and performance aspects of using noVNC, SPICE, and Apache Guacamole as solutions for providing access to the desktop via a browser. One focus of the paper is the redirection of the sound interface of the virtual desktop. The paper's authors conclude that Guacamole, in combination with the Remote Desktop Protocol (RDP), is the best solution. The authors evaluated the lag time for remote audio playback in an Internet scenario. by measuring the time between the automated initiation of the live streaming using the media player software VLC and the arrival of the music data by analyzing network traffic via monitoring software Wireshark. When using Guacamole combined with the protocol RDP, the average lag was around 750\,ms. When using Guacamole combined with the protocol Virtual Network Computing (VNC), the average lag was around 1750\,ms. The paper also includes a short evaluation of the lag for video frames represented by background color changes of a Java application. Again, the configuration Guacamole and RDP offered the best performance video update with an average lag of around 300\,ms. The configuration of Guacamole and VNC caused an average lag of around 500\,ms.

Magana et al.~\cite{magana2019remote} compared in 2019 the most popular remote desktop protocols and software implementations. The authors analyzed five protocols: PCoIP used in the Amazon WorkSpaces, RDP, TeamViewer, VNC~\cite{rfc6143}, and Citrix Independent Computing Architecture (ICA). In the paper, the network transfer rate and its relation to the quality experienced by the DaaS user are evaluated by assuming three scenarios: using an office software suite, web browsing, and video streaming. One emphasis of the paper is comparing the transfer rate of the different protocols for three scenarios. In the Office Software Suite scenario, the transfer rate is measured and compared when performing several different tasks like opening and editing a document (typing), loading an image, and saving a document. The web browsing scenario covers requesting three different web pages, and for the video streaming scenario, the same YouTube video file was viewed at different resolutions. The authors conclude that RDP requires fewer downstream and upstream network resources for all three scenarios. It is important to understand that the protocols RDP and VNC are implemented in open-source projects and proprietary solutions for many different operating systems. Depending on the different quality levels of the many existing implementations, it is difficult to generalize about the performance of the protocols, as individual implementations may have better or worse performance characteristics than expected.

@@

For both cloud service categories, users expect a latency value that is not considered annoying when interacting with the service and its applications. Choy et al.~\cite{ChoyWongSimonRosenberg2012}, Claypool et al.~\cite{claypool2010latency}, and Jarschel et al.~\cite{jarschel2011evaluation} consider a maximum tolerable latency, based on subjective tests, being around 100\,ms. This relatively low value sets the demand for a fast network interaction between client and service and limits their maximum distance. As mentioned in Clincy et al.~\cite{clincy2013subjective}, The former cloud gaming service provider OnLive specified in the year 2010 that the distance between service provider and consumer should not be greater than 1000 miles (approx. 1600\,km) because otherwise the round-trip communications delay time between will be too long for video games.

@@

The open-source server virtualization platform Proxmox VE runs bare-metal and offers virtual machines with the Kernel-based Virtual Machine (KVM) and container-based virtualization via LXC Linux container. Since the API of LXC is, in comparison to the Docker API, rather poorly documented and lacks several features, we considered Docker a better container virtualization solution. A deployment of docker directly in the host-operating system or inside a virtual machine is both possible.

>>>>

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%
\begin{document}
%
\title{Challenges and Solutions of Developing and
	Implementing a Revolutionary Novel Desktop-as-a-Service}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Christian Baun \and 
Johannes Bouché}
% \author{Christian Baun\orcidID{0009-0004-9955-3752} \and
% 	Johannes Bouché\orcidID{TODO}}
%
\authorrunning{C. Baun, J. Bouché}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Faculty of Computer Science and Engineering,\\
Frankfurt University of Applied Sciences,\\
Nibelungenplatz 1, 60318 Frankfurt am Main, Germany
\email{[christianbaun|johannes.bouche]@fb2.fra-uas.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
Desktop-as-a-Service (DaaS) is a cloud service category with great potential for improving the everyday work of running and using applications. However, existing DaaS solutions have several drawbacks, which have limited the popularity of this service category in the past. Our novel DaaS solution called DESIGN not only overcomes the limitations of existing solutions but also offers significant benefits. It has the potential to revolutionize the way we use applications, supporting non-modified Linux and Windows applications with just a browser on the client side. This paper describes some of the knowledge we gained during the development because of the challenges we faced and the solutions we found.

@@

This paper discusses challenges during the development and implementation of a novel Desktop-as-a-Service (DaaS) solution called DESIGN that enables the deployment and usage of unmodified Linux and Windows applications in the same way as web applications and can run inside public resources but can also be deployed in a private context. Since all interaction with DESIGN and the applications that reside inside is done via a user's web browser, the users can use any client, no matter what hardware or host operating system it is compromised of. A browser is the only software component required to use the novel DaaS service.

@@

This document is structured as follows. Section~\ref{sec:relatedworkAbschnitt} discusses related work on DaaS solutions from an academic perspective, whereas section~\ref{sec:RecommendDaaSarchitecture} describes the state-of-the-art DaaS architecture DESIGN we designed and implemented. Section~\ref{sec:AnalysisPossibleComponents} presents the most relevant challenges we faced, analyses our options for handling these, and describes the solution we found. Finally, section~\ref{sec:Conclusions} discusses conclusions, and section~\ref{sec:NextSteps} includes directions for future work.

@@

Celesti et al.~\cite{celesti2016improving} implemented in 2016 a DaaS using the IaaS solution OpenStack~\cite{sefraoui2012openstack} and analyzed the characteristics and performance aspects of using noVNC, which is a client for the protocol Virtual Network Computing (VNC) that is implemented in the Hypertext Markup Language (HTML) and JavaScript, the protocol SPICE (Simple Protocol for Independent Computing Environments), and the remote desktop gateway Apache Guacamole as solutions for providing access to the desktop via a browser. One focus of the paper is the redirection of the sound interface of the virtual desktop. The paper's authors conclude that Guacamole, in combination with the Remote Desktop Protocol (RDP), is the best solution. The authors evaluated the lag time for remote audio playback in an Internet scenario. by measuring the time between the automated initiation of the live streaming using the media player software VLC and the arrival of the music data by analyzing network traffic via monitoring software Wireshark. When using Guacamole combined with the protocol RDP, the average lag was around 750\,ms. When using Guacamole combined with the protocol VNC, the average lag was around 1750\,ms. The paper also includes a short evaluation of the lag for video frames represented by background color changes of a Java application. Again, the configuration Guacamole and RDP offered the best performance video update with an average lag of around 300\,ms. The configuration of Guacamole and VNC caused an average lag of around 500\,ms.

Magana et al.~\cite{magana2019remote} compared in 2019 the most popular remote desktop protocols and software implementations. The authors analyzed five protocols: PCoIP used in the Amazon WorkSpaces, RDP, TeamViewer, VNC, and Citrix Independent Computing Architecture (ICA). In the paper, the network transfer rate and its relation to the quality experienced by the DaaS user are evaluated by assuming three scenarios: using an office software suite, web browsing, and video streaming. One emphasis of the paper is comparing the transfer rate of the different protocols for three scenarios. In the Office Software Suite scenario, the transfer rate is measured and compared when performing several different tasks like opening and editing a document (typing), loading an image, and saving a document. The web browsing scenario covers requesting three different web pages, and for the video streaming scenario, the same YouTube video file was viewed at different resolutions. The authors conclude that RDP requires fewer downstream and upstream network resources for all three scenarios. It is important to understand that the protocols RDP and VNC are implemented in open-source projects and proprietary solutions for many different operating systems. Depending on the different quality levels of the many existing implementations, it is difficult to generalize about the performance of the protocols, as individual implementations may have better or worse performance characteristics than expected.

@@

For every deployment scenario or cloud service category possible, users expect a latency value that is not annoying when interacting with the service and its applications. Choy et al.~\cite{ChoyWongSimonRosenberg2012}, Claypool et al.~\cite{claypool2010latency}, and Jarschel et al.~\cite{jarschel2011evaluation} consider a maximum tolerable latency, based on subjective tests, being around 100\,ms. This relatively low value sets the demand for a fast network interaction between client and service and limits their maximum distance. As mentioned in Clincy et al.~\cite{clincy2013subjective}, The former cloud gaming service provider OnLive specified in the year 2010 that the distance between the service provider and the consumer should not be greater than 1000 miles (approx. 1600\,km) because otherwise, the round-trip communications delay time between will be too long for video games.

@@

The open-source server virtualization platform Proxmox VE runs bare-metal and offers virtual machines with the Kernel-based Virtual Machine (KVM) and container-based virtualization via LXC Linux container. Since the LXC API is, in comparison to the Docker API, rather poorly documented and lacks several features, we considered Docker a better container virtualization solution. A deployment of docker directly in the host-operating system or inside a virtual machine is both possible.
